---
title: "Midterm Project"
author: "Chuning Yuan"
date: "10/19/2019"
output: pdf_document
---

# The Question

Is there a correlation between GDP per Capita and prevalence of HIV in the 15–49 age bracket? And if yes, how strong is that correlation? How about the correlation between GDP per Capita and prevalence of HIV in female?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(WDI)
library(tidyverse)
library(kableExtra)
```


# The Data
In this data analysis, I use the data available on World Bank webpage.Here are a brief description of each data set:

**GDP/capita** : GDP per capita is gross domestic product divided by midyear population. GDP is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products.  Data are in current U.S. dollars.  

**Prevalence of HIV among people aged 15–49 (%)**: Prevalence of HIV refers to the percentage of people ages 15-49 who are infected with HIV.  

**Prevalence of HIV, female among people aged 15-24(%)**: Prevalence of HIV, female is the percentage of females who are infected with HIV. Youth rates are as a percentage of the relevant age group.  

The following code shows how I download the dataset:
```{r}
#getdata
new_wdi_cache <- WDIcache()
WDIsearch("Prevalence of HIV, total")
WDIsearch("Prevalence of HIV, female")
WDIsearch("gdp.*capita.*US\\$", cache = new_wdi_cache)


```



# Combing dataframe & data cleaning

\qquad We remove all entries that are aggregated regional values and then we rename the indicators.Then we combine the three dataframes to allow us to compare GDP per capita and HIV prevalence and HIV prevalence, female. 
```{r}
#cleandata
wdi_data <- WDI(indicator = c("NY.GDP.PCAP.CD","SH.DYN.AIDS.ZS","SH.HIV.1524.FE.ZS"), start = 1960, end = 2018, extra = TRUE)
names(wdi_data)
wdi_data <- subset(wdi_data, region != "Aggregates")
names(wdi_data)[which(names(wdi_data) == "NY.GDP.PCAP.CD")] <- "GDP"
names(wdi_data)[which(names(wdi_data) == "SH.DYN.AIDS.ZS")] <- "HIV_total"
names(wdi_data)[which(names(wdi_data) == "SH.HIV.1524.FE.ZS")] <- "HIV_female" 

data=na.omit(wdi_data)
names(data)
```

\qquad Datasets often feature missing data.So we need to take a  look at the percentage of missing data in the combined dataframe.


\qquad About 24.3% of the GDP per Capita column in the combined dataframe have missing data. This is quite substantial and is most likely due the fact that consistent measurements of GDP are costly and have only started in the last few decades according to the World Bank webpage (It was found in 1945 after WWII).
 And we have an  at 68.5% for HIV Prevalence missing data and 69.1% for HIV Prevalence of female in the combined dataframe. The lag in consistent measurements of HIV associated metrics have only really been performed on a large scale from the early-1980s when HIV/Aids became a recognised major health crisis.

# Observed the distribution of each data sets

```{r}
p1=ggplot(data, aes(x = GDP) ) +
  geom_histogram(bins = 100,fill="darkblue")
summary(data$GDP)

p2=ggplot(data, aes(x = HIV_total)) +
  geom_histogram(bins = 20,fill="darkblue")
summary(data$HIV_total)

p3=ggplot(data, aes(x = HIV_female)) +
  geom_histogram(bins = 20,fill="darkblue")
summary(data$HIV_female)

gridExtra::grid.arrange(p1,p2,p3,ncol= 2)
```

\qquad The results above shows that some countries in our dataframe are considerable pulling the distribution’s mean up (compared to the median). This is particularly the case with the HIV Prevalence data.


\qquad Big picture of the data
```{r}
#data transformation
growth <- data %>% group_by(year) %>%
summarize(HIV_total=mean(HIV_total), HIV_female=mean(HIV_female), GDP=mean(GDP))


kable(growth, digits = 4, align = "c",booktabs=TRUE ,caption = "WorldwideTrend ",col.names = c("year", "HIV_female","HIV_total","GDP" ))%>% kable_styling(latex_options = "HOLD_position")
```

\qquad The table above listed all the averge GDP and average HIV prevalence in total and female in each year. From the table we can see the basic trend of the GDP growth and the precentage of the prevalence of HIV  growth from 1990-2018. After 2009 the prevalence HIV in total and in felmale has goes down slightly each year while the GDP are still keep growing,in my opinion, this might be the new invention of the hpv immunizations begin to be popular.




# The Plots
## Scatter plots
I generate various plots to get an overview of the distribution and attempt to identify trends and patterns.
First, we look at the overall data set and generate a scatter plot of GDP per Capita and HIV Prevalence in for the 136 countries listed in our dataset from 1990 to 2018.
```{r warning=F,message=F}
#scatter plot of GDP per Capita
ggplot(data, aes(x = year, y = GDP), title = "GDP")+
  geom_point(size = 1,color="lightgreen") + scale_x_continuous("Year",breaks = seq(1990,2018,5)) + geom_smooth(color="blue",size=0.5) 

#scatter plot of Prevalence of HIV total
ggplot(data, aes(x = year, y = HIV_total), title = "HIV_total")+
  geom_point(size = 1,color="darkred") + scale_x_continuous("Year",breaks = seq(1990,2018,5)) + geom_smooth(color="blue",size=0.5)  

#scatter plot of Prevalence of HIV female
ggplot(data, aes(x = year, y = HIV_female), title = "HIV_female")+
  geom_point(size = 1,color="lightblue") + scale_x_continuous("Year",breaks = seq(1990,2018,5)) + geom_smooth(color="blue",size=0.5)  

```
\qquad According to our data, the rate of HIV prevalence has slight increased between 1990 and 2010 with a stagnation in the mean from the early 2000s and slight decline since 2005. This would correspond to advances in preventative measures to reduce the incidence and likelihood of contracting HIV.






# Correlation and Linear Regression

\qquad The scatter plot clearly indicates that the lower GDP per Capita data points (i.e. countries) have a much higher HIV prevalence compared countries with higher GDP per Capita.

\qquad Let’s take a closer look by creating a plot with a square root scale applied to the x-axis to further emphasise countries with lower GDP per capita. We’ll also use the R function geom_smooth() to perform a simple linear regression to better visualise the relationship between the two variables.
```{r}
#plot HIV Prevalance vs. GDP per Capita.
ggplot(data, aes(x = GDP, y = HIV_female)) + geom_point(size = 1L,alpha=0.5,colour = "#0c4c8a") +geom_smooth(method = 'gam',color="red",size=0.5) 

# square root scale of GDP
ggplot(data) +
 aes(x = GDP, y = HIV_total) +
 geom_point(size = 1L, colour = "#0c4c8a",alpha=0.5) +
 scale_x_continuous(trans = "sqrt") +
 theme_minimal()+geom_smooth(method = 'gam',color="red",size=0.5)


```
The scatter plot above further indicates that countries which smaller GDP per capita have on average higher HIV prevalence.


```{r}
#plot HIV Prevalance female vs. GDP per Capita.
p4=ggplot(data, aes(x = GDP, y = HIV_female)) + geom_point(size = 1L,alpha=0.5,colour = "#0c4c8a") +geom_smooth(method = 'gam',color="red",size=0.5)  

# square root scale of GDP
p5=ggplot(data) +
 aes(x = GDP, y = HIV_female) +
 geom_point(size = 1L, colour = "#0c4c8a",alpha=0.5) +
 scale_x_continuous(trans = "sqrt") +
 theme_minimal()+geom_smooth(method = 'gam',color="red",size=0.5)


```

# Explore more of the data
\qquad Futhermore, I want to select top ten HIV prevalance in total and in female countries for the most rencently year(2018) of the data set. So I made a table to arrange the data first, then I compared the GDP of those ten countries in 2018 with their HIV to see if their is any correlation.
```{r}
#select the top ten HIV Prevalance countries in most recent years
da = data %>% filter(year == 2018) %>% dplyr::select (country,HIV_total,HIV_female)%>%arrange(desc(HIV_total))
kable(head(da,10),align = "c")%>% kable_styling(latex_options = "HOLD_position")

#select the top ten HIV Prevalance of female countries in most recent years
da = data %>% filter(year == 2018) %>% dplyr::select (country,HIV_total,HIV_female)%>%arrange(desc(HIV_female))
kable(head(da,10),align = "c")%>% kable_styling(latex_options = "HOLD_position")
```

```{r}

sub_HIV_total <- subset(data, country %in% c("Eswatini", "Lesotho", "South Africa",
"Botswana", "Zimbabwe", "Mozambique", "Namibia", "Zambia", "Malawi", "Equatorial Guinea"))
ggplot(subset(sub_HIV_total, year == 2018), aes(x = GDP, y = HIV_total, color = country)) +geom_point() +geom_text(aes(label = country), size=3, nudge_y = 0.4) +
scale_x_continuous(limits = c(0, 15000))


sub_HIV_female <- subset(data, country %in% c("Eswatini", "Lesotho", "South Africa",
"Botswana", "Zimbabwe", "Mozambique", "Namibia", "Zambia", "Malawi", "Equatorial Guinea"))
ggplot(subset(sub_HIV_female, year == 2018), aes(x = GDP, y = HIV_female, color = country )) +geom_point() +geom_text(aes(label = country), size=3, nudge_y = 0.4) +
scale_x_continuous(limits = c(0, 15000))

```


\qquad From the plots above I can't tell wether those two variable has linear relationship because some countries who has higher prevalance HIV rate also have higher GDP than some other country (ex. Eswatini), and Equatorial Guinea which has the lowest prevalance HIV rate in these ten countries but also have the highest GDP among them.Therefore this is not consistent with what I have found before. We may consider other factor at this point such as countries enviornment, geogarphical location and their development level of the medical facilities.


# Conclusion

\qquad In this project, I collected data from public sources (World Bank). And I did an initial exploratory data analysis. Then, I derived a correlation factor and applied linear regression to assess the linear relationship between three interest (GDP per capita, HIV prevalence, HIV prevalence in female). In addition, I also try to see if the current top ten HIV prevalence countries have any representative information to support what I found, but the results turns out not so helpful, which is normal.Therefore, we may need futher analysis on this topic.


